---
title: 'Supporting Information: Contact angle statistics'
author: "Clifton Anderson"
date: "July 2, 2018"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(xlsx)
library(car)
```

## Plots and ANOVA results

A model was fitted to consider the effects of (1) Drop location, (2) Liquid, and (3) the plasma treatment.

```{r Data importing and manipulating}

CAraw <- read.xlsx2('G:/Documents/MS-Thesis-Pitt/Data-analysis/contact angles.xlsx',sheetIndex=1,colClasses = NA)
CAraw$Date <- as.Date(CAraw$Date)
printed <- CAraw[CAraw$Finish!='PP',]
ggplot(CAraw,aes(x=Treatment,y=avg,color=Liquid))+geom_point()+facet_grid(Finish~.)

# Duplicate treatment for more descriptive factor name
CAraw$CoatingType <- CAraw$Treatment 

# Collapse the factor levels into the type of treatment
levels(CAraw$CoatingType) <- list(plasma=c('blown3min.1', 'blown3min.2',
                                         'quick.dry', 'plas_coated'),
                                clean= c('as is', 'clean', 'etoh_ace', 'washed',
                                         'wiped', 'wiped_etoh_ace', 'wiped_water'),
                                paa=   c('paa'),
                                support=c('support'))

# Remove unwanted factor levels
TreatmentSum <- filter(CAraw,!Treatment %in% c('as is',  'paa', 'support', 'washed', 'wiped_water') &
                         !Finish %in% c('glass') &
                          Time=='SS' &
                         !Liquid %in% c('.1plasma','.33plasma','.67plasma')) %>%
  dplyr::select(.,Finish,Liquid,CoatingType,Treatment,avg) %>%
  droplevels()

# summarize group means 
CAtable <- TreatmentSum %>% group_by(Finish,Liquid,Treatment) %>%
  summarise(n=n(),
            Contact.Angle=mean(avg),
            SD = sd(avg),
            'Cos(theta)'=cos(3.14159/180*Contact.Angle)) #%>%
  # knitr::kable(.,
               # caption='contact angle of water and plasma on clean and plasma coated surfaces',
               # digits=c(0,0,0,0,1,1,3))


CAtable
# convert CA table to html for easy pasting into word.
# htmlTable(txtRound(CAtable[,1:ncol(CAtable)],digits=c(0,0,1,0,1,1,3)))

```


### Dilution experiments
```{r dilution experiments}
# Get data for dilution experiments

dil <- filter(CAraw,Date=='2019-06-18') %>% droplevels()
levels(dil$Liquid) <- c(.1,.33,.67,0,1,10)
dil$Liquid <- as.numeric(as.character(dil$Liquid))

ggplot(dil,aes(Liquid,avg,color=Treatment,shape=Finish))+
  geom_jitter(width=0.01)+
  scale_shape_manual(values=c(19,1))+
  xlab('Volume fraction plasma')+
  ylab('Contact angle')+
  theme_minimal()
  # ggtitle('Contact angle in plasma-PBS dilutions')

# continuous dilution levels
dil.lm.a <- lm(avg~Liquid,data=filter(dil,Liquid>0, Finish=='glossy', Treatment=='quick.dry'));summary(dil.lm.a)
dil.lm.i <- lm(avg~Treatment*Liquid,data=dil[dil$Liquid>0,]);summary(dil.lm.i)
Anova(dil.lm.a,type=2)

# Categorical dilution levels
dil.lm.c <- lm(avg~Treatment+factor(Liquid)+Finish,data=dil[dil$Liquid>0,]);summary(dil.lm.c)

# test for lack of fit:
anova(dil.lm.c,dil.lm.a) # no sign of lack of fit

```


**Analyses exclude pure PBS**
Treating dilutions as categorical, Liquid is significant at p=0.01
Treating dilutions on continuous scale, Liquid is significant (from ANOVA, type=2) at p=0.0020
Lack of fit p=0.5128, so I can't say that a linear relationship is a bad one.



## Plasma coating t-tests

```{r plasma coating t-tests}
SS <- subset(CAraw,Time=='SS' & Finish=='glossy' & Treatment != 'paa')
SS %>% 
  filter(.,Treatment!='clean'& Treatment != 'wiped',Liquid=='plasma') %>%
  pairwise.t.test(.$avg,.$Treatment,data=.,pool.sd=TRUE,p.adj='bonf')
```

```{r compare quick.dry matte vs glossy}
QD <- CAraw %>% filter(Treatment=='quick.dry',Finish!='PP',Liquid=='plasma') %>% droplevels()
with(QD,table(Finish,Liquid))
t.test(avg~Finish,data=QD,var.equal=TRUE)
ggplot(QD,aes(sample=avg))+stat_qq(aes(color=Finish))
```


```{r PP coating tests}
PP <- CAraw[CAraw$Finish=='PP',]
PP %>% group_by(Treatment,Liquid) %>% summarise(n=n(),Angle=mean(avg),SD=sd(avg),Conf95=SD/sqrt(n)*qt(0.95,df=n-1),'Cos'=cos(Angle*3.14159/180))
ggplot(PP,aes(Treatment,avg,color=Liquid)) + geom_jitter(width=0.25)

```




