---
title: "Factorial analysis"
output:
  html_document: default
  html_notebook: default
---

# $3^2$ factorial experiments analysis

```{r load libraries and import data, message=FALSE, warning=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(xlsx)
library(car)
library(RcmdrMisc)
library(pwr)
```

```{r import and process data}
raw <- read.xlsx("G:/Documents/MS-Thesis-Pitt/Data-analysis/Spinning.xlsx",
                 sheetIndex = 2,
                 colClasses = NA)

raw <- filter(raw,bac_recovery != is.na(bac_recovery)) #bac_recovery>0.2)
raw$BR <- raw$bac_recovery*100
raw$Vol_out2in <- 1-raw$HCT_diluted/100/.85
raw$CR_trap <- raw$conc_recovery*raw$Vol_out2in

raw$Hold_time.num <- raw$Hold_time
raw$HCT_diluted.num <- raw$HCT_diluted
raw$Hold_time <- factor(x=raw$Hold_time)#,ordered=TRUE)
raw$HCT_diluted <- factor(x=raw$HCT_diluted)#,ordered=TRUE)


# calculate PR and RR <- pasted from "Spinning analysis.Rmd"
Temp = 23 #C
rho_PBS <- -5.2037e-6*Temp^2 - 8.1463e-6*Temp + 1.0097 
# I got this correlation from
# (http://www.sciencedirect.com/science/article/pii/S003991400400373X)

raw$VolFracPlasma  <- with(raw,{VolWB*(1-HCT)/(VolWB*(1-HCT) + VolPBS)})
raw$RhoSolution    <- with(raw,{(rho_PBS*VolPBS + 1.024*(1-HCT)*VolWB)/(VolPBS + (1-HCT)*VolWB)})
raw$VFlowdown      <- with(raw,{gplas_tot/(RhoSolution*(1-HCT_plas)+1.093*HCT_plas)})

raw$RR <- with(raw,{VFlowdown*HCT_plas/(VolWB*HCT)*100})
raw$PR <- with(raw,{VFlowdown*(1-HCT_plas)/(VolPBS+VolWB*(1-HCT))*100})

# subset data (with or without outliers from linear models. Make sure to justify each remove in comments)
inliers <- raw[-c(11,39,41,21),]

# summarize results
raw.sum <- inliers %>% 
  group_by(HCT_diluted,Hold_time)  %>%
  summarise(n=n(),
            BR.mean = mean(BR),BR.sd = sd(BR),
            PR.mean = mean(PR),PR.sd = sd(PR),
            RR.mean = mean(RR),RR.sd = sd(RR))
raw.sum
```


```{r Plotting bacteria recoveries}
ggplot(raw,aes(HCT_diluted,BR,color=Hold_time,shape=Donor))+geom_jitter(width=0.05)+ylim(ymin=0,ymax=100)
with(raw,plotMeans(BR,HCT_diluted,Hold_time,error.bars='conf.int'))

ggplot(inliers,aes(Hold_time,conc_recovery,color=Donor))+geom_point()+facet_grid(.~HCT_diluted)
ggplot(raw,aes(Hold_time.num,conc_recovery/Vol_out2in,color=Donor))+geom_point()+facet_grid(.~HCT_diluted)
ggplot(inliers,aes(Hold_time,conc_recovery*Vol_out2in/0.85))+geom_point()+facet_grid(.~HCT_diluted)
```

```{r plot red cell recovery}
ggplot(inliers,aes(HCT_diluted,RR,color=Hold_time,shape=Donor))+geom_jitter(width=0.05)
```
Red cell recovery decreases with   as the cell pack compresses

```{r plot plasma recovery}
ggplot(inliers,aes(HCT_diluted.num,PR,color=Hold_time,shape=Donor))+geom_jitter(width=1)
ggplot(inliers,aes(HCT_diluted.num,VFlowdown,color=Hold_time,shape=Donor))+geom_jitter(width=1)
ggplot(inliers,aes(VFlowdown,PR,color=Hold_time,shape=Donor))+geom_jitter(width=1)+facet_grid(.~HCT_diluted)
```
At H=5, time doesn't seem to matter. As hematocrit increases, Plasma recovery increases with time because the cell pack compresses more.

Mashsa saw that plasma recovery was higher with short times. Potential reasons:
> Matte disks do not dry as quickly
> 

At H=47, Plasma is undiluted, which increases the cost of trapped plasma.
I need to revisit trough overflow for red cells. 

#### Is it wettability or viscosity?
 If wettability: Measure contact angle for dilutions of plasma-PBS
 If Viscosity: interpolate plas and PBS visc.


```{r Linear models for BR}
options(contrasts = c('contr.sum','contr.sum'))
BR.lm.i <- lm(BR~Hold_time*HCT_diluted+Donor,data=inliers); summary(BR.lm.i)
BR.lm.a <- lm(BR~Hold_time+HCT_diluted+Donor,data=raw); summary(BR.lm.a)
Anova(BR.lm.i,type=2)
```


```{r BR Diagnostics}
pwr.f2.test(u=13,sig.level=0.01,f2=0.64/(1-0.64),power=0.99)

termplot(BR.lm.i,partial.resid = TRUE,ask=FALSE,col.res =c('green','blue','red','black','pink','purple')[raw$Donor],ylim = c(-25,25))
with(BR.lm.i,plot(fitted.values,residuals,col=c('green','blue','red')[as.numeric(raw$Hold_time)]))
```
## Commentary
From partial residual plots, the donor makes a big difference.



```{r Concentration recovery models}
ggplot(inliers,aes(Hold_time.num,conc_recovery,color=HCT_diluted))+geom_point()+geom_smooth(method='lm',se=F)
ggplot(inliers,aes(HCT_diluted.num,conc_recovery,color=Hold_time))+geom_point()+geom_smooth(method='lm',se=F)


CR.lm.poly  <- lm(conc_recovery~Hold_time*HCT_diluted+Donor,
             data=inliers,
             contrasts=list(Hold_time='contr.poly',
                            HCT_diluted='contr.poly',
                            Donor='contr.sum'))
summary(CR.lm.poly)
# I need to either research polynomial contrasts more, or else convert time and HCT to continuous.
# maybe I need to center these. How can I look at VIF?
CR.lm.num <- lm(conc_recovery~Donor+Hold_time.num*HCT_diluted.num,data=raw)
summary(CR.lm.num)
Anova(CR.lm.num,type=2)
```

Significant positive linear interaction between hold time and hematocrit.

for facet_grid(Hold_Time~.):
At low hold time,  increasing the hematocrit decreases the concentration recovery.
At high hold time, increasing the hematocrit increases the concentration recovery.

At low hematocrit, increasing hold time decreases concentration recovery
At high hematocrit,increasing hold time increases concentration recovery
 > More plasma backflow





```{r Plot everything}
pairs(dplyr::select(raw,c(Hold_time,HCT_diluted,BR,conc_recovery,VFlowdown,PR,RR)),
      col = c('blue','green','red')[inliers$HCT_diluted])
```




```{r setting up data frame and plotting results}
HCT = factor(c(1,0,1,-1,0,-1,-1,0,1),ordered=TRUE)
time= factor(c(0,-1,-1,-1,1,0,1,0,1),ordered=TRUE)
Bac_sc = c(35,48,33,61,42,46,57,43,32)
Bac_mc = c(35,44,35,63,41,44,61,38,35)
df <- data.frame(HCT=HCT,
                 time=time,
                 Bac_sc=Bac_sc,
                 Bac_mc=Bac_mc)

ggplot(df,aes(HCT,Bac_sc,color=time))+geom_point()

BR.lm <- lm(Bac_sc ~ time+HCT,data=df);summary(BR.lm)
```
