---
title: "Hematocrit"
author: "Clifton Anderson"
date: "May 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hematocrit comparisons between clinical and out centrifuge
Clinical centrifuge: 10,000xg for 5 minutes
Our centrifuge: 1300xg for 10 minutes

Variables:
> HCT: nominal hematocrit based on assumption of gravity packing = 80%
> clin.1: 1st spin in clinical centrifuge
> clin.2: 2nd spin in clinical centrifuge
> Pitt.1: 1 spin in Pitt centrifuge

Female blood, 1 day old. Some red cells were congealed on pipetting.


```{r Import and graph}
library(ggplot2)
library(dplyr)

# Load table
df_raw <- read.table('clinical_calibration_long.csv',header=TRUE,sep=',')

# Plot results
ggplot(df_raw,aes(HCT,HCT.meas,color=Centrifuge))+
  geom_jitter(width=0.05)
  # geom_smooth(method='lm',formula=y~x,se=FALSE)

df <- df_raw %>% 
  #filter(Centrifuge!='clin.1') %>% 
  droplevels()

# summarize group means
means <- df %>% 
          group_by(HCT,Centrifuge) %>% 
          summarise(n=n(),
                    HCT.mean=mean(HCT.meas),
                    sd=sd(HCT.meas)) %>%
          summarise(ratio=HCT.mean)
          as.data.frame()

# Set up linear model with interactions
df$HCT <- as.factor(df$HCT)
H.lm <- lm(HCT.meas ~ HCT*Centrifuge, data=df) ;summary(H.lm)
```

### Results
A linear model says that Pitt overpredicts HCT by 1.642 (even with interactions). The interaction values are excess differences between Pitt and clin.2 (i.e. 1.6417-1.0750 = 0.5666. The Pitt centrifuge overestimated hematocrit by 0.5666 at H~20.)

I want to make a more usable correlation between Pitt and clinical hematocrits


```{r wide data}
df_wide <- read.table('clinical_calibration.csv',header=TRUE,sep=',')
df_wide

ggplot(df_wide,aes(clin.2,Pitt.1))+geom_point()

H.lm <- lm(Pitt.1~clin.2,data=filter(df_wide,HCT>=20));summary(H.lm)
H.lm <- lm(Pitt.1~clin.2,data=filter(df_wide,HCT>20));summary(H.lm)


```
This model artificially pairs data, but the slope is useful. It says that Pitt hematocrit overpredicts true hematocrit by 5% +- 0.5

I took ratios of group means in excel. see clinical_calibration_wide.csv